---
title: "Predicting Exercise Quality Using Accelerometer Data"
author: "Jim Novy"
date: "November 22, 2015"
output: html_document
---

#Summary

In this paper we will see a technique for building a model to predict exercise quality using accelerometer data.  We will then use it to predict some unlabeled data.

#Data Load

Here we will load the testing and training dataset.  We will then remove all columns that are either not represented in the testing set or that cause inferential errors.

```{r, cache=TRUE}
library(dismo)
library(doParallel)
library(caret)
#Set seed
set.seed(1)
#Read in data
testing <- read.csv("pml-testing.csv")
training <- read.csv("pml-training.csv")
#Remove values without predictive data in testing set 
testing <- testing[,!names(testing) %in% c("X","raw_timestamp_part_1","raw_timestamp_part_2","cvtd_timestamp","new_window","kurtosis_roll_belt","kurtosis_picth_belt","kurtosis_yaw_belt","skewness_roll_belt","skewness_roll_belt.1","skewness_yaw_belt","max_roll_belt","max_picth_belt","max_yaw_belt","min_roll_belt","min_pitch_belt","min_yaw_belt","amplitude_roll_belt","amplitude_pitch_belt","amplitude_yaw_belt","var_total_accel_belt","avg_roll_belt","stddev_roll_belt","var_roll_belt","avg_pitch_belt","stddev_pitch_belt","var_pitch_belt","avg_yaw_belt","stddev_yaw_belt","var_yaw_belt","var_accel_arm","avg_roll_arm","stddev_roll_arm","var_roll_arm","avg_pitch_arm","stddev_pitch_arm","var_pitch_arm","avg_yaw_arm","stddev_yaw_arm","var_yaw_arm","kurtosis_roll_arm","kurtosis_picth_arm","kurtosis_yaw_arm","skewness_roll_arm","skewness_pitch_arm","skewness_yaw_arm","max_roll_arm","max_picth_arm","max_yaw_arm","min_roll_arm","min_pitch_arm","min_yaw_arm","amplitude_roll_arm","amplitude_pitch_arm","amplitude_yaw_arm","kurtosis_roll_dumbbell","kurtosis_picth_dumbbell","kurtosis_yaw_dumbbell","skewness_roll_dumbbell","skewness_pitch_dumbbell","skewness_yaw_dumbbell","max_roll_dumbbell","max_picth_dumbbell","max_yaw_dumbbell","min_roll_dumbbell","min_pitch_dumbbell","min_yaw_dumbbell","amplitude_roll_dumbbell","amplitude_pitch_dumbbell","amplitude_yaw_dumbbell","var_accel_dumbbell","avg_roll_dumbbell","stddev_roll_dumbbell","var_roll_dumbbell","avg_pitch_dumbbell","stddev_pitch_dumbbell","var_pitch_dumbbell","avg_yaw_dumbbell","stddev_yaw_dumbbell","var_yaw_dumbbell","kurtosis_roll_forearm","kurtosis_picth_forearm","kurtosis_yaw_forearm","skewness_roll_forearm","skewness_pitch_forearm","skewness_yaw_forearm","max_roll_forearm","max_picth_forearm","max_yaw_forearm","min_roll_forearm","min_pitch_forearm","min_yaw_forearm","amplitude_roll_forearm","amplitude_pitch_forearm","amplitude_yaw_forearm","var_accel_forearm","avg_roll_forearm","stddev_roll_forearm","var_roll_forearm","avg_pitch_forearm","stddev_pitch_forearm","var_pitch_forearm","avg_yaw_forearm","stddev_yaw_forearm","var_yaw_forearm")]
training <- training[,!names(training) %in% c("X","raw_timestamp_part_1","raw_timestamp_part_2","cvtd_timestamp","new_window","kurtosis_roll_belt","kurtosis_picth_belt","kurtosis_yaw_belt","skewness_roll_belt","skewness_roll_belt.1","skewness_yaw_belt","max_roll_belt","max_picth_belt","max_yaw_belt","min_roll_belt","min_pitch_belt","min_yaw_belt","amplitude_roll_belt","amplitude_pitch_belt","amplitude_yaw_belt","var_total_accel_belt","avg_roll_belt","stddev_roll_belt","var_roll_belt","avg_pitch_belt","stddev_pitch_belt","var_pitch_belt","avg_yaw_belt","stddev_yaw_belt","var_yaw_belt","var_accel_arm","avg_roll_arm","stddev_roll_arm","var_roll_arm","avg_pitch_arm","stddev_pitch_arm","var_pitch_arm","avg_yaw_arm","stddev_yaw_arm","var_yaw_arm","kurtosis_roll_arm","kurtosis_picth_arm","kurtosis_yaw_arm","skewness_roll_arm","skewness_pitch_arm","skewness_yaw_arm","max_roll_arm","max_picth_arm","max_yaw_arm","min_roll_arm","min_pitch_arm","min_yaw_arm","amplitude_roll_arm","amplitude_pitch_arm","amplitude_yaw_arm","kurtosis_roll_dumbbell","kurtosis_picth_dumbbell","kurtosis_yaw_dumbbell","skewness_roll_dumbbell","skewness_pitch_dumbbell","skewness_yaw_dumbbell","max_roll_dumbbell","max_picth_dumbbell","max_yaw_dumbbell","min_roll_dumbbell","min_pitch_dumbbell","min_yaw_dumbbell","amplitude_roll_dumbbell","amplitude_pitch_dumbbell","amplitude_yaw_dumbbell","var_accel_dumbbell","avg_roll_dumbbell","stddev_roll_dumbbell","var_roll_dumbbell","avg_pitch_dumbbell","stddev_pitch_dumbbell","var_pitch_dumbbell","avg_yaw_dumbbell","stddev_yaw_dumbbell","var_yaw_dumbbell","kurtosis_roll_forearm","kurtosis_picth_forearm","kurtosis_yaw_forearm","skewness_roll_forearm","skewness_pitch_forearm","skewness_yaw_forearm","max_roll_forearm","max_picth_forearm","max_yaw_forearm","min_roll_forearm","min_pitch_forearm","min_yaw_forearm","amplitude_roll_forearm","amplitude_pitch_forearm","amplitude_yaw_forearm","var_accel_forearm","avg_roll_forearm","stddev_roll_forearm","var_roll_forearm","avg_pitch_forearm","stddev_pitch_forearm","var_pitch_forearm","avg_yaw_forearm","stddev_yaw_forearm","var_yaw_forearm")]
```

Next we will use dismo's kfolds function to generate stratified samples.  This allows us to ensure that each fold has similar representative amounts of samples from each class we're predicting.  We then use these folds to generate our training and testing sets:

```{r, cache=TRUE}
folds <- kfold(training,6,by=training$classe)
#Subset training and testing data
trainmdl <- training[folds %in% c(2:6),]
testmdl <- training[folds %in% 1,]
```

Here we instruct the doParallel library to utilize all cores in the system minus one, then train a RandomForest model with the data constraining the control to 3 folds and produce the confusion matrix on out of sample predictions:
```{r, cache=TRUE}
c1 <- makeCluster(detectCores()-1)
registerDoParallel(c1)
model <- train(classe~.,trainmdl,method='rf',trControl = trainControl(method = "cv", number = 3))
testmdlnoclasse <- testmdl[,!names(testmdl) %in% c("classe")]
confusionMatrix(testmdl$classe,predict(model,testmdlnoclasse))
```

Here we see that our out of sample prediction is 99.7% accurate.  In running the same method using the exact parameters but a different seed and / or a different index, the worst performance seen during initial analysis was 99.6%, which inside the bound for the CI in this sample.

Finally we load in the instructor's pml_write_files function, predict on the unlabeled set and write that to files for submission:

```{r, cache=TRUE}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
predict(model,testing)
pml_write_files(predict(model,testing))

```
